---
title: "Flexible polyline encoding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flexible polyline encoding}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Flexible polyline encoding

The `flexpolyline` R package provides a binding to the
[C++ implementation](https://github.com/heremaps/flexible-polyline/tree/master/cpp) of the
flexible polyline encoding by [HERE](https://github.com/heremaps/flexible-polyline).
The flexible polyline encoding is a lossy compressed representation of a list of
coordinate pairs or coordinate triples.

The encoding is achieved by:

(1) Reducing the decimal digits of each value;
(2) encoding only the offset from the previous point;
(3) using variable length for each coordinate delta; and
(4) using 64 URL-safe characters to display the result.
The felxible polyline encoding is a variant of the [Encoded Polyline Algorithm Format](https://developers.google.com/maps/documentation/utilities/polylinealgorithm) by Google.

**Notes:**

* Decoding gives reliable results up to a precision of 7 digits.
The tests are also limited to this range.
* The order of the coordinates (lng, lat) does not correspond to the original C ++ implementation (lat, lng).
This enables simple conversion to `sf` objects, without reordering the columns.
* The encoding is lossy, this means the encoding process could reduce the precision of your data.


## C++ binding

The package contains two C++ function, which bind directly to the C++
implementation of the flexible polyline encoding. These functions target a tight
interface from C++ to R and have the same input and output as their counterpart
in the original implementation. In order to achieve the maximum performance
(e.g. in a package), these should be used.


## Encode
Encoding a line is straight forward by passing a matrix (column order is
longitude, latitude) with coordinates to `encode()`. The default precision value
is `5`. If a matrix with two columns is provided, the third dimension is
automatically set to `"ABSENT"`:

```{r encode2d}
library(flexpolyline)

line2d <- matrix(
  c(8.69821, 50.10228,
    8.69567, 50.10201,
    8.69150, 50.10063,
    8.68752, 50.09878),
  ncol = 2, byrow = TRUE
)

(encodes2d <- encode(line2d, precision = 5))
```

If a matrix with three columns is passed a line with a third dimension is encoded.
The type of the third dimension can be passed to the encoding and includes the
following selection, that are specified via an integer value in the `third_dim` argument:

* 0: ABSENT
* 1: LEVEL
* 2: ALTITUDE
* 3: ELEVATION (default)
* 4: RESERVED1 (not available)
* 5: RESERVED2 (not available)
* 6: CUSTOM1
* 7: CUSTOM2

By default the precision of the third dimension `third_dim_precision` is set to
the same value as the precision value for the longitude and latitude dimensions.
```{r encode3d}
line3d <- matrix(
  c(8.69821, 50.10228, 10.11111,
    8.69567, 50.10201, 20.22222,
    8.69150, 50.10063, 30.33333,
    8.68752, 50.09878, 40.44444),
  ncol = 3, byrow = TRUE
)

(encodes3d <- encode(line3d, precision = 5, third_dim = 3, third_dim_precision = 5))
```


## Decode

In order to decode the lines from above, the encoded line is passed to the `decode()` function.
The function detects the from the encoding if it is a two dimensional or three dimensional line:
```{r decode2d}
decode(encodes2d)
```

In the case of a three dimensional line, the third column is named after the name
of the third dimension that was chosen during the encoding (in this example
`"ELEVATION"`):
```{r decode3d}
decode(encodes3d)
```


## References

* [Flexible polyline encoding by HERE](https://github.com/heremaps/flexible-polyline)
* [Encoded Polyline Algorithm Format](https://developers.google.com/maps/documentation/utilities/polylinealgorithm)
* Inspired by the [googlePolylines](https://github.com/SymbolixAU/googlePolylines) package
