---
title: "Simple feature support"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple feature support}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Simple features

A common way to deal with spatial data in R is using the [sf](https://cran.r-project.org/web/packages/sf/index.html)  package, which is
built on the concept of simple features. According to the the sf package [vignette](https://r-spatial.github.io/sf/articles/sf1.html),
simple features are *"...a formal standard (ISO 19125-1:2004) that describes how
objects in the real world can be represented in computers, with emphasis on the
spatial geometry of these objects. It also describes how such objects can be
stored in and retrieved from databases, and which geometrical operations should
be defined for them."*

The most common simple feature geometry types are: POINT, LINESTRING, POLYGON,
MULTIPOINT, MULTILINESTRING, MULTIPOLYGON. Whereby all geometries are based on POINTs.
In the flexible polyline encoding binding in this package only LINESTRINGs are supported

There are four possible dimension combinations of simple features package.
In the `flexpolyine` package the first three dimension combinations are be represented
as follows:

* XY: Represented as line with LNG and LAT dimensions.
* XYz: Represented as line with LNG, LAT and LEVEL/ALTITUDE/ELEVATION as third dimension.
* XYM: Represented as aline with LNG, LAT and CUSTOM1/CUSTOM2 as third dimension.
* XYZM: Not supported by the flexible polyline encoding.

## Encoding and decoding
### Simple feature geometry (sfg)
If no value for `third_dim` is provided in `encode_sf()`, the dimensions of the
sf geometry are considered. In case of `"XYZ"` the third dimension is set to
`"ELEVATION"` and thereby recognized again as `"XYZ"` by `decode_sf()` during
decoding:
```{r sfg_z}
library(flexpolyline)
library(sf)

coords <- matrix(
    c(8.69821, 50.10228, 10.11111,
      8.69567, 50.10201, 20.22222,
      8.69150, 50.10063, 30.33333,
      8.68752, 50.09878, 40.44444),
    ncol = 3, byrow = TRUE
)

(sfg_z <- st_linestring(coords, dim = "XYZ"))

(sfg_enc_z <- encode_sf(sfg_z))

decode_sf(sfg_enc_z)
```

In case of `"XYM"` the third dimension is set to `"CUSTOM1"` and again set to `"XYM"` in decoding:
```{r sfg_m}
(sfg_m <- st_linestring(coords, dim = "XYM"))

(sfg_enc_m <- encode_sf(sfg_m))

decode_sf(sfg_enc_m)
```

### Simple feature geometry list-column (sfc)
The geometry column in simple feature data sets is called simple feature geometry
list-column and is supported as input in encoding:
```{r sfc}
(sfc <- st_as_sfc(
  lapply(seq(1, 5), function(x) {
    st_linestring(coords[, 1:2] + runif(1, -1, 1), dim = "XY")
  }),
  crs = 4326
))

(sfc_enc <- encode_sf(sfc))

decode_sf(sfc_enc)
```

### Simple feature (sf)
A simple feature is a geometry that is connected with further characteristics
(other than it's coordinates), which in practice means a `data.frame`
(also `data.table` or `tibble`) with an `sfc` column. The `flexpolyline` package
only takes care of the coordinates of the LINESTRINGs in the sf object, other
columns are not encoded and stored in the string:
```{r sf}
(sf <- st_as_sf(
  data.frame(
    id = seq(1,length(sfc)),
    color = sample(c("red", "green", "blue"), 5, replace = TRUE),
    geometry = sfc
  )
))

(sf_enc <- encode_sf(sf))

decode_sf(sf_enc)
```
**Note:** The columns `"id"` and `"color"` were not encoded and are missing
after decoding.

## References

* [CRAN: Simple Features for R](https://cran.r-project.org/web/packages/sf/index.html)
* [Wikipedia: Simple Features](https://en.wikipedia.org/wiki/Simple_Features)
* [OGC: Simple Feature Access](https://www.ogc.org/standards/sfa)
